---
title: "Intro to RR (Demo)"
---

Reproducible research
===

The workflow of reproducible research implies that we use programming code to carry out certain tasks in a scientific research project:
1. Load and inspect data  
2. Draw statistical graphics  
3. Fit statistical models  
4. Compare observed and modeled data  
5. Produce report that describes all this  

A. Load and inspect data
``` {r chucnk1}
ds <- mtcars
head(ds, 10)
```

B. Draw statistical graphics
``` {r chuck, echo=FALSE, message=FALSE}
require(ggplot2)
ggplot(data=mtcars, aes(x=wt, y=mpg)) +geom_point()
```

C. Fit statistical model
``` {r chunk3}
lm(formula = mpg ~ wt, mtcars)
```

D. Comparing observed and modeled data
``` {r chunk4}
ds <- mtcars 
modelA <- lm(formula = mpg ~ wt, mtcars)
ds$mpg_modelA <- predict(modelA)
head(ds)
p <- ggplot(data=ds, aes(x=wt, y=mpg))
p <- p + geom_point()
p <- p + geom_line(aes(y=mpg_modelA), color ="red")
p
```

F. Produce report that describes all this 

Conclusion: as we see the heavier the car is, the less mpg it gets