---
title: "lmer guide"
output:
  html_document:
    css: ~/GitHub/Longitudinal_Models_of_Religiosity_NLSY97/www/css/thesis.css
    fig.retina: 2
    fig_width: 8
    theme: united
    toc: yes
  md_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: no
    fig_crop: no
    fig_width: 8
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    fig_width: 6.5
mainfont: Calibri
---

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
require(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.

```

```{r set_options, echo=F, message=F}
require(knitr)
# getwd()
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
#   fig.height = 4.8, 
#   fig.width = 6.5, 
#   out.width = NULL,
  fig.path = 'lmerGuide/',     
  dev = "png",
  dpi = 100
)
echoChunks <- FALSE
warningChunks<- FALSE
options(width=120) #So the output is 50% wider than the default.
read_chunk("./Models/LCM/LCM.R") # the file to which knitr calls for the chunks
```

```{r DeclareGlobals, echo=echoChunks, message=FALSE}
aesDefs
```

```{r LoadPackages, echo=echoChunks, message=F}
```

```{r LoadData, echo=echoChunks, message=T}
# select only respondence in the cross-sectional sample

```

Locating model estimation results in S4 objects produced by of <code>lmer()</code> calls of *lme4* package.
[back to Appendix][appendix]


## Basics 
Prepare data for modeling. 
Only the first 200 respondents will be selected to keep illustration light.
```{r}
dsL<-readRDS("./Data/Derived/dsL.rds")
ds<- dsL %>% dplyr::filter(id %in% c(1:200),year %in% c(2000:2011)) %>% 
  dplyr::mutate(timec=year-2000, timec2= timec^2, timec3= timec^3, 
                agec= round( (agemon/12),0)-16) %>% 
  dplyr::select(id,year,attend, timec,timec2, timec3, agec)
head(ds, 20)
```

Fit the model with <code>lmer</code>
```{r}
m10 <-lmer (attend ~ 
               1  + agec + timec + timec2 + timec3
             + agec:timec +agec:timec2 + agec:timec3
             + (1 + timec + timec2 + timec3 | id),
             data = ds, REML=0)
model<- m10
```

Print the basic results of  the fitted model
```{r}
print(model) 
```

Or get a more detailed summary
```{r}
summary(model)
```

For a list of objects that can be extracted from the model
```{r}
str(summary(model))
```



There a number of ways one could extract the needed element from the S4 object. In addition, some of the elements might be stored in several locations.   

## Model formula
```{r}
model@call # 1
(summary(model))$call #2
```

## Fit and Information indices

```{r}
# get indicies
mInfo<-summary(model)$AICtab
mInfo
mInfo['logLik']
```

Aleternatively,
```{r}
logLik<-logLik(model)
dev<-deviance(model)
AIC <- AIC(model) 
BIC <- BIC(model)
N<- model@devcomp$dims["N"] # Looks like the total number of data points ( individual-by-time)
p<- model@devcomp$dims["p"] # Looks like the number of estimated parameters, verify
ids<- (summary(model))$ngrps # number of units on level-2, here: individuals
mInfo<- c("logLik"=logLik,"dev"=dev,"AIC"=AIC,"BIC"=BIC,"N"=N, "p"=p, "ids"=ids)
mInfo
```

## Random Effects (RE)

### Matrix of RE
```{r}
str(summary(model)$varcor)
# extract RE covariance matrix
mRE<- data.frame(summary(model)$varcor$id) 
mRE
mRE<- data.frame(VarCorr(model)$id)
mRE
```

```{r}
mREcor<-  data.frame(attr(summary(model)$varcor$id,"correlation")) # corrleation 
mREcor
mREcor<- data.frame( attr(VarCorr(model)$id,"correlation"))
mREcor
mREsd<-   data.frame(STD=(attr(summary(model)$varcor$id,"stddev"))) 
mREsd
```



### extracting RE for each individual
```{r}
RE<- lme4:::ranef.merMod(model)$id 
head(RE,5)
tail(RE,5)
mRE
cor(RE)  # not the same as mRE, find out why
var(RE)  # not the same as mRE, find out why
```

## Fixed Effects (RE)

### Matrix of FE
```{r}
str(summary(model)$vcov)
mFE<- (summary(model)$vcov@factors$correlation) # notice that this is object of class corMatrix
str(mFE)
```

### estimate of the FE
```{r}
# similar ways to extract FE estimates, #3 is the fullest
FE1<- fixef(model) #1 
FE2<- getME(model, "beta") # 2
FE3<- summary(model)$coefficients # 3
coefs<- FE1
```

## Prediction and Residuals
### Restatement of model input
First, recover information that went into the model
```{r}
# dsP - P for predicted
dsp<-data.frame(model@frame) # original vars used by the model (no interaction terms)
head(dsp,13)
```

Another way, which includes interaction terms, but no outcome
```{r}
dsp<- data.frame(getME(model,"X")) # no Y, only predictors (with interaction terms)
head(dsp,13)
```

We can add response vector and the grouping factor for the second level (individual)
```{r}
dsp$id<-getME(model,"flist")$id # first level grouping factor, individual
dsp$y<-getME(model,"y") # response vector
head(dsp,13)
```

### Adding model output
There are several way to extract the predictions made by the model
```{r}
# model outcome, predicted values
yHat1<- fitted(model) # 1
yHat2<- predict(model) # 2
yHat3<- getME(model,"mu") # 3

identical(yHat1,yHat2)
identical(as.numeric(yHat2),yHat3)
dsp$yHat<- predict(model)
head(dsp,13)
```

### Adding residual
```{r}
dsp$resid<- lme4:::residuals.merMod(model) # individual residual (id and time)
head(dsp,13)
identical (  dsp$y-dsp$yHat, dsp$resid)

```

Getting the standard error of residuals
```{r}
getME(model,"sigma") # standard error of residuals, same sigma(model)
sigma(model) # std.error of residuals <- this methods is preferred
# however
sd(dsp$resid) # not the same as sigma(model) = find out why
identical (sigma(model),sd(dsp$resid)) # WHY?
```

## Conditional values
Predictions form fixed effects only, no individual variability is used in calculation

```{r}
# create object "coefs" for easy reference
coefs <- fixef(model)
# use fixed effects estimates to find conditional predictions??
dsp$yPar<-(
  (coefs["(Intercept)"])         +(coefs["agec"]*dsp$agec)
  +(coefs["timec"]*dsp$timec)    +(coefs["agec:timec"]*dsp$agec*dsp$timec)
  +(coefs["timec2"]*dsp$timec2)  +(coefs["agec:timec2"]*dsp$agec*dsp$timec2)
  +(coefs["timec3"]*dsp$timec3)  +(coefs["agec:timec3"]*dsp$agec*dsp$timec3)
)
head(dsp,13)

```

```{r child, child = '../../Models/Descriptives/Citations.Rmd'}

```


